Estimating Correspondences of Deformable Objects ``In-the-wild"

### 2015-11-05 contents/implementation.tex

\noindent\textbf{Step 4:} Utilise the dense correspondences acquired by the optical flow in order to automatically generate dense and sparse (on the outline) landmark annotations and build a dense \cite{ramnath2008increasing, Amberg2009, anderson2014using} and part-based \cite{Tzimiropoulos2014} AAM, respectively.

@@

{\refstepcounter{steps}\label{sec:step4}\subsection*{Step 4: Dense and Part-Based Active Appearance Models}}
% \subsection{Dense and Part-Based Active Appearance Models}

The deformation fields obtained from Step~\ref{sec:step3} can be used to naturally build two different kinds of effective AAMs \cite{Cootes2001, Matthews2004}: \emph{dense} \cite{ramnath2008increasing, Amberg2009, anderson2014using} and \emph{part-based} \cite{Tzimiropoulos2014}. The only difference between these two AAM formulations is on the way that the shape is represented and, thus, the manner in which the texture is sampled. Each one of them is suitable for object classes with specific properties. The dense AAM provides an exceptionally effective modeling and fitting for non-articulated objects, such as ears and faces, whose appearance has characteristic structures that spread all over their region (even if these structures cannot be consistently annotated). On the other hand, there exist other challenging object classes, such as arms and legs, that not only cannot be consistently annotated with landmarks, but their appearance is distinctive only on the object's outline and not in its interior region. Especially in the case of human body parts, they are almost always covered by clothes, which makes it impossible to construct robust texture models.

@@

\paragraph{Outline Part-Based Active Appearance Model}
For

As already mentioned, for estimating body pose and modeling other object types with similar challenges, we propose to only consider the established correspondences on the object outline.

Based on the proposed shape flow framework, we develop a methodology for robust automatic landmarking of the object outline. This allows a small initial training set that is manually annotated without consistent landmarks to be expanded to a much larger training set in an automatic way. In more detail, regarding the manually annotated images, for which SVS functions have been constructed, the outline correspondences are yielded by evaluating the estimated shape flow on the object outline of the reference image. To expand the training set to images without any annotations, we fit the trained dense AAM on these images and then only consider the deformed positions of the dense landmarks that correspond to the object outline on the mean shape.

Having reliable object outline annotations on a large training set, we appropriately fuse their information to infer consistent landmarks on the object skeleton. This procedure equips deformable model training with consistent landmark annotations. This is particularly valuable, since it automates the training process and overcomes the limitations of inconsistent and inaccurate manual annotations on challenging objects such as forearms and legs.

Based on these skeleton landmarks, we then train an parts-based deformable model. More specifically, ...

>>>>

\noindent\textbf{Step 4:} Utilise the dense correspondences acquired by the optical flow in order to automatically generate dense and sparse (on the outline) landmark annotations and build a dense \cite{ramnath2008increasing, Amberg2009, anderson2014using} and patch-based \cite{Tzimiropoulos2014} AAM, respectively.

@@

{\refstepcounter{steps}\label{sec:step4}\subsection*{Step 4: Dense and Patch-Based Deformable Models}}
% \subsection{Dense and Patch-Based Active Appearance Models}

The deformation fields obtained from Step~\ref{sec:step3} can be used to naturally build two different kinds of effective Active Appearance Models (AAMs) \cite{Cootes2001, Matthews2004}: \emph{dense} \cite{ramnath2008increasing, Amberg2009, anderson2014using} and \emph{patch-based} \cite{Tzimiropoulos2014}. The only difference between these two AAM formulations is on the way that the shape is represented and, thus, the manner in which the texture is sampled. Each one of them is suitable for object classes with specific properties. The dense AAM provides an exceptionally effective modeling and fitting for non-articulated objects, such as ears and faces, whose appearance has characteristic structures that spread all over their region (even if these structures cannot be consistently annotated). On the other hand, there exist other challenging object classes, such as arms and legs, that not only cannot be consistently annotated with landmarks, but their appearance is distinctive only on the object's outline and not in its interior region. Especially in the case of human body parts, they are almost always covered by clothes, which makes it impossible to construct robust texture models.

@@

\paragraph{Outline Patch-Based Active Appearance Model}
The object classes for which the interior appearance does not have specific structure are modeled using patch-based AAMs \cite{Tzimiropoulos2014} trained on a set of sparse landmarks. Especially for human body parts (arms, legs), we strongly believe that the points located to the outline of the object are more suitable compared to the internal ones that correspond to the skeleton joints, which are commonly used by current literature \cite{buehler2011upper,charles2013domain,pfister2015flowing,yang2013articulated}.

The main differences between the patch-based and dense AAMs are that (a)~the densified shape instances are sub-sampled to include only the outline points, and (b)~the texture representation involves the sampling a neighbourhood around each point instead of a single pixel. Specifically, in order to build the outline sparse shape model, we simply select the outline points on the SVS reference frame. Then, by taking advantage of the dense correspondences obtained by Step \ref{sec:step3}, the shape model is trained in a similar way as in the dense case. Moreover, similar to the dense case, the texture model is built by sampling the image values from the sparse shape locations, i.e. $\mathcal{W}(\bm{s}(\bm{p}))$. However, contrary to dense AAMs, we sample a patch that is centred around each landmark point. These patches are then vectorised and concatenated in a single texture vector. Note that the optimisation process remains exactly the same.

### 2015-11-05 contents/introduction.tex

Furthermore, we capitalise on recent advances on multiframe optical flow estimation \cite{Garg:2013hu} and show that the relevant methodologies have matured enough to densely annotate the proposed shapes using either simplistic or even more sophisticated and robust shape representation methods \cite{Nguyen2013}.
In particular, in order to build dense correspondences between different shape instances of the same object class, we jointly estimate the optical flow among all the instances by imposing low-rank constrains, an approach that we call \emph{Shape Flow}. Multiframe optical flow has originally been applied on video sequences, relying on the assumptions of color consistency and motion smoothness \cite{Garg:2013hu}. However, these assumptions do not hold in our case, where we have a collection of shapes. Therefore, we introduce appropriate modifications based on the consistency of image-based shape representation, as well as low rank priors.
%
%On the other hand,  while shapes having no motion smoothness completely as all training data are independent. Low rank constrain plays important role to simulate motion smoothness assumption.

In addition, we demonstrate how dense Active Appearance Models (AAMs) can be built after establishing dense shape correspondences. We show that the proposed methodology can be applied on landmark localisation, even though it is not tailored for that task, achieving particularly good performance. We also illustrate the performance of the proposed methodology for describing objects that do not have consistent landmarks. Furthermore, we show how to build a powerful part-based AAM on the sparse outline landmarks that are obtained by sampling the shape of the dense AAM. Using the resulting outline part-based AAM, we report state-of-the-art performance on the task of human body parts localisation on challenging databases. Finally, we show that the proposed part-based AAM can be used to provide consistent annotations for different body parts.

@@

  \item We report state-of-the-art quantitative and qualitative results on the task of human body parts localisation by employing a part-based AAM trained on the outline landmarks that are sampled by the dense AAM. Our proposed model outperforms all current state-of-the-art techniques that are trained on skeleton joints

  \item We show that the employed part-based AAM corrects the annotations that are currently provided for most major human body pose databases. Upon acceptance of the paper, we will make the corrected annotations publicly available.

>>>>

Furthermore, we capitalise on recent advances on multiframe optical flow estimation \cite{Garg:2013hu} and show that the relevant methodologies have matured enough to densely annotate the proposed shapes using either simplistic or even more sophisticated and robust shape representation methods \cite{Nguyen2013}.
In particular, in order to build dense correspondences between different shape instances of the same object class, we jointly estimate the optical flow among all the instances by imposing low-rank constrains, an approach that we call \emph{Shape Flow}. Multiframe optical flow has originally been applied on video sequences, relying on the assumptions of colour consistency and motion smoothness \cite{Garg:2013hu}. However, these assumptions do not hold in our case, where we have a collection of shapes. Therefore, we introduce appropriate modifications based on the consistency of image-based shape representation, as well as low rank priors.
%
%On the other hand,  while shapes having no motion smoothness completely as all training data are independent. Low rank constrain plays important role to simulate motion smoothness assumption.

Additionally, we show that the proposed methodology can be applied on landmark localisation, even though it is not tailored for that task, achieving particularly good performance. Specifically, we explain how to build powerful dense SDMs that are suitable for objects that have rich interior texture but lack landmarks consistency. Furthermore, we show how to build a powerful patch-based SDMs on the sparse outline landmarks of objects that do not have semantically meaningful interior textures. Using the resulting outline patch-based SDM, we report state-of-the-art performance on the task of human body parts localisation on challenging databases. Finally, we show that the proposed patch-based SDM can be used to provide consistent annotations for different body parts.

@@

  \item We report state-of-the-art quantitative and qualitative results on the task of human body parts localisation by employing a patch-based SDM trained on the outline landmarks that are sampled by the dense correspondences. Our proposed model outperforms all current state-of-the-art techniques that are trained on skeleton joints

  \item We show that the employed patch-based SDM corrects the annotations that are currently provided for most major human body pose databases. Upon acceptance of the paper, we will make the corrected annotations publicly available.
